---
title: "Final Projects; Winter Institute in Data Science"
author: "Adam Weber, Ali Amini, John Burzawa"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
## Load packages and import two data set

```{r, echo=TRUE}
library(devtools) 
library(dplyr)
library(here)
library(tidyverse)
library(haven)
library(ggplot2)


anes_2020 <- read.csv(here("anes_2020.csv"))
anes_2016 <- read_dta("~/anes_timeseries_2016.dta")
```
## cleaning data
```{r , echo=FALSE}
# Rename variable  for free trade ------------------------------

anes_2020 <- anes_2020 %>% rename(
  party_id_orig = V201018,
  free_trade = V202359, 
  free_trade_intensity = V202360, 
  free_trade_relationships = V202362,
  free_trade_job = V202365, 
  free_trade_intl = V202367x)


anes_2016<- anes_2016 %>% rename(party_id_orig = V161019,
                                 free_trade = V162176, 
                                 free_trade_intensity = V162176a)



# Create party ID variable ------------------------------

anes_2020 <- anes_2020 %>%  mutate( 
  pid_chr = case_when(
    party_id_orig == 2 ~ "rep",
    party_id_orig == 1 ~ "dem",
    party_id_orig == 4 ~ "ind"),
  republican = if_else(pid_chr == "rep", "yes", "no"))

anes_2016 <- anes_2016 %>%  mutate(
  pid_chr = case_when(
    party_id_orig == 2 ~ "rep",
    party_id_orig == 1 ~ "dem",
    party_id_orig == 4 ~ "ind"),
  republican = if_else(pid_chr == "rep", "yes", "no"))

```


## In the following steps we will clean data and create approprite varibles to show diffrences in mean and standard deviation(sd) for each, we use sd as apporximate for polrization; we do sampe porcess for free trade, health care, persident job, and wall at Mexico, we will include all codes for the free trdae and we will do the same process for others and only the output and t test will be shown.

#FAVOR OR OPPOSE FREE TRADE AGREEMENTS WITH OTHER COUNTRIES
```{r, echo=TRUE}
anes_2020 <- anes_2020 %>% 
  mutate (opinion_free_trade = case_when( free_trade == 3 ~ 0,
                              free_trade == 1 & free_trade_intensity == 1 ~ 3,  
                             free_trade == 1 &  free_trade_intensity == 2 ~ 2,
                              free_trade == 1 &  free_trade_intensity == 3 ~ 1,
                                free_trade == 2 &  free_trade_intensity == 1 ~ -3,
                               free_trade == 2 &  free_trade_intensity == 2 ~ -2,
                              free_trade == 2 &  free_trade_intensity == 3 ~ -1))  

anes_2016 <- anes_2016 %>% 
  mutate(opinion_free_trade = case_when( free_trade == 3 ~ 0,
                             free_trade == 1 & free_trade_intensity == 1 ~ 3,  
                             free_trade == 1 &  free_trade_intensity == 2 ~ 2,
                             free_trade == 1 &  free_trade_intensity == 3 ~ 1,
                             free_trade == 2 &  free_trade_intensity == 1 ~ -3,
                             free_trade == 2 &  free_trade_intensity == 2 ~ -2,
                             free_trade == 2 &  free_trade_intensity == 3 ~ -1))



#calculation of Standard Deviation and mean for each party in free  trade----
anes_2020 %>% group_by(pid_chr) %>%
  summarise(sd_free_trade = sd(opinion_free_trade, na.rm = TRUE))

anes_2020 %>% group_by(pid_chr) %>%
  summarise(mean_free_trade = mean(opinion_free_trade, na.rm = TRUE))



anes_2016 %>% group_by(pid_chr) %>%
  summarise(sd_free_trade = sd(opinion_free_trade, na.rm = TRUE))

anes_2016 %>% group_by(pid_chr) %>% 
  summarise(mean_free_trade = mean(opinion_free_trade, na.rm = TRUE))
# t-test----
t.test (anes_2016$opinion_free_trade, anes_2020$opinion_free_trade)

#visualization----
ggplot(data = anes_2016,
       mapping = aes(x=anes_2016$opinion_free_trade, fill = anes_2016$pid_chr)) +
  geom_bar(position = "dodge") + scale_fill_manual(values=c('blue', 'green', 'red', 'white')) +theme_dark()

ggplot(data = anes_2020,
       mapping = aes(x=anes_2020$opinion_free_trade, fill = anes_2020$pid_chr)) +
  geom_bar(position = "dodge") + scale_fill_manual(values=c('blue', 'green', 'red', 'white')) +theme_dark()

boxplot(anes_2016$opinion_free_trade, anes_2020$opinion_free_trade,
        names = c('2016', '2020'), col=c('red', 'blue'), main='Public oppinion in free trade')


ggplot(anes_2016, aes(anes_2016$pid_chr,anes_2016$opinion_free_trade)) + 
  geom_boxplot(col=c('blue','green', 'red', 'white'))

ggplot(anes_2020, aes(anes_2020$pid_chr,anes_2020$opinion_free_trade)) + 
  geom_boxplot(col=c('blue','green', 'red', 'white'))


```



#INCREASE/DECREASE GOVERNMENT SPENDING TO HELP PAY FOR HEALTH CARE

```{r , echo=FALSE}
anes_2020 <- anes_2020 %>% 
  mutate (opinion_health_care = case_when( V202380x == 4 ~ 0,
                                         V202380x == 1  ~ 3,  
                                         V202380x == 2  ~ 2,
                                         V202380x == 3  ~ 1,
                                         V202380x == 7  ~ -3,
                                         V202380x == 6  ~ -2,
                                         V202380x == 5 ~ -1)) 

anes_2016 <- anes_2016 %>% 
  mutate(opinion_health_care= case_when( V162193x == 4 ~ 0,
                                         V162193x == 1  ~ 3,  
                                         V162193x == 2  ~ 2,
                                         V162193x == 3  ~ 1,
                                         V162193x == 7  ~ -3,
                                         V162193x == 6  ~ -2,
                                         V162193x == 5 ~ -1)) 


#calculation of Standard Deviation and mean for each party in health care----

anes_2020 %>% group_by(pid_chr) %>%
  summarise(sd_health_care = sd(opinion_health_care, na.rm = TRUE))

anes_2020 %>% group_by(pid_chr) %>%
  summarise(mean_health_care = mean(opinion_health_care, na.rm = TRUE))


anes_2016 %>% group_by(pid_chr) %>%
  summarise(sd_health_care = sd(opinion_health_care, na.rm = TRUE))

anes_2016 %>% group_by(pid_chr) %>%
  summarise(mean_health_care = mean(opinion_health_care, na.rm = TRUE))


# t-test-----
t.test(anes_2016$opinion_health_care, anes_2020$opinion_health_care)


#visualization----


ggplot(data = anes_2016,
       mapping = aes(x=anes_2016$opinion_health_care, fill = anes_2016$pid_chr), col= c('blue', 'green', 'red','white')) +
  geom_bar(position = "dodge")+ scale_fill_manual(values=c('blue', 'green', 'red', 'white')) +theme_dark()

ggplot(data = anes_2020,
       mapping = aes(x=anes_2020$opinion_health_care, fill = anes_2020$pid_chr)) +
  geom_bar(position = "dodge")+ scale_fill_manual(values=c('blue', 'green', 'red', 'white')) +theme_dark()


boxplot(anes_2016$opinion_health_care,anes_2020$opinion_health_care, 
        names = c('2016', '2020'), col=c('red', 'blue'), main='Public oppinion in health care')


ggplot(anes_2016, aes(anes_2016$pid_chr,anes_2016$opinion_health_care)) + 
  geom_boxplot(col=c('blue','green', 'red', 'white'))

ggplot(anes_2020, aes(anes_2020$pid_chr,anes_2020$opinion_health_care)) + 
  geom_boxplot(col=c('blue','green', 'red', 'white'))

```

#FAVOR OR OPPOSE BUILDING A WALL ON BORDER WITH MEXICO

```{r, echo=FALSE}
# create a variable for building war at Mexico borders-------------
anes_2020 <- anes_2020 %>% 
  mutate (opinion_wall_Mexico = case_when( V201426x == 4 ~ 0,
                                           V201426x == 1  ~ 3,  
                                           V201426x == 2  ~ 2,
                                           V201426x == 3  ~ 1,
                                           V201426x == 7  ~ -3,
                                           V201426x == 6  ~ -2,
                                           V201426x == 5 ~ -1)) 


anes_2016 <- anes_2016 %>% 
  mutate (opinion_wall_Mexico = case_when( V161196x == 4 ~ 0,
                                           V161196x == 1  ~ 3,  
                                           V161196x == 2  ~ 2,
                                           V161196x == 3  ~ 1,
                                           V161196x == 7  ~ -3,
                                           V161196x == 6  ~ -2,
                                           V161196x == 5 ~ -1))


#calculation of Standard Deviation and mean for each party in  building war at Mexico -------

anes_2020 %>% group_by(pid_chr) %>%
  summarise(sd_wall_Mexico  = sd(opinion_wall_Mexico , na.rm = TRUE))


anes_2020 %>% group_by(pid_chr) %>%
  summarise(mean_wall_Mexico  = mean(opinion_wall_Mexico , na.rm = TRUE))


anes_2016 %>% group_by(pid_chr) %>%
  summarise(sd_wall_Mexico  = sd(opinion_wall_Mexico , na.rm = TRUE))


anes_2016 %>% group_by(pid_chr) %>%
  summarise(mean_wall_Mexico  = mean(opinion_wall_Mexico , na.rm = TRUE))

# t.test----------
t.test(anes_2016$opinion_wall_Mexico, anes_2020$opinion_wall_Mexico)


# visualization---------

ggplot(data = anes_2016,
       mapping = aes(x=anes_2016$opinion_wall_Mexico, fill = anes_2016$pid_chr)) +
  geom_bar(position = "dodge")+scale_fill_manual(values=c('blue', 'green', 'red', 'white')) +theme_dark()

ggplot(data = anes_2020,
       mapping = aes(x=anes_2020$opinion_wall_Mexico, fill = anes_2020$pid_chr)) +
  geom_bar(position = "dodge") +scale_fill_manual(values=c('blue', 'green', 'red', 'white')) +theme_dark()


boxplot(anes_2016$opinion_wall_Mexico,anes_2020$opinion_wall_Mexico, 
        names = c('2016', '2020'), col=c('red', 'blue'), main='Public oppinion in wall on Mexico')


ggplot(anes_2016, aes(anes_2016$pid_chr,anes_2016$opinion_wall_Mexico)) + 
  geom_boxplot(col=c('blue','green', 'red', 'white'))

ggplot(anes_2020, aes(anes_2020$pid_chr,anes_2020$opinion_wall_Mexico)) + 
  geom_boxplot(col=c('blue','green', 'red', 'white'))


```

#APPROVE OR DISAPPROVE PRESIDENT HANDLING JOB
```{r, echo=FALSE}
# create a variable for PRESIDENT HANDLING JOB scale -2, -1, +1, +2-------------------

anes_2020 <- anes_2020 %>% 
  mutate (opinion_persident_job = case_when(V201129x == 3 ~ -1,
                                            V201129x == 4 ~ -2,
                                            V201129x == 2  ~ 1,
                                            V201129x == 1  ~ 2)) 



anes_2016 <- anes_2016 %>% 
  mutate (opinion_persident_job = case_when(V161082x == 3 ~ -1,
                                            V161082x == 4 ~ -2,
                                            V161082x == 2  ~ 1,
                                            V161082x == 1  ~ 2))

#calculation of Standard Deviation and mean for PRESIDENT HANDLING JOB -------



anes_2020 %>% group_by(pid_chr) %>%
  summarise(sd_opinion_persident_job = sd(opinion_persident_job, na.rm = TRUE))

anes_2020 %>% group_by(pid_chr) %>%
  summarise(mean_opinion_persident_job = mean(opinion_persident_job, na.rm = TRUE))


anes_2016 %>% group_by(pid_chr) %>%
  summarise(sd_opinion_persident_job = sd(opinion_persident_job, na.rm = TRUE))

anes_2016 %>% group_by(pid_chr) %>%
  summarise(mean_opinion_persident_job = mean(opinion_persident_job, na.rm = TRUE))

# t.test-------------
t.test(anes_2016$opinion_persident_job, anes_2020$opinion_persident_job)




# visualization--------------

ggplot(data = anes_2016,
       mapping = aes(x=anes_2016$opinion_persident_job, fill = anes_2016$pid_chr)) +
  geom_bar(position = "dodge")+scale_fill_manual(values=c('blue', 'green', 'red', 'white')) +theme_dark()

ggplot(data = anes_2020,
       mapping = aes(x=anes_2020$opinion_persident_job, fill = anes_2020$pid_chr)) +
  geom_bar(position = "dodge")+scale_fill_manual(values=c('blue', 'green', 'red', 'white')) +theme_dark()

boxplot(anes_2016$opinion_persident_job,anes_2020$opinion_persident_job, 
        names = c('2016', '2020'), col=c('red', 'blue'), main='Public oppinion in persident job ')


ggplot(anes_2016, aes(anes_2016$pid_chr,anes_2016$opinion_persident_job)) + 
  geom_boxplot(col=c('blue','green', 'red', 'white'))

ggplot(anes_2020, aes(anes_2020$pid_chr,anes_2020$opinion_persident_job)) + 
  geom_boxplot(col=c('blue','green', 'red', 'white'))



```

#GOVERNMENT ACTION ABOUT RISING TEMPERATURES

```{r, echo=FALSE}
  # create a variable for rising temperatures-------------

anes_2020 <- anes_2020 %>% 
  mutate (opinion_rising_temperatures = case_when( V201401 == 3    ~ 0,
                                                   V201401 == 1 & V201402 == 1 ~ 3,  
                                                   V201401 == 1 & V201402 == 2 ~ 2,
                                                   V201401 == 1  & V201402 == 3 ~ 1,
                                                   V201401 == 2 & V201402 == 1 ~ -3,
                                                   V201401 == 2 &V201402 == 2 ~ -2,
                                                   V201401 == 2 & V201402 == 3 ~ -1))  
anes_2016 <- anes_2016 %>% 
  mutate (opinion_rising_temperatures = case_when( V161225x == 4 ~ 0,
                                                   V161225x == 1  ~ 3,  
                                                   V161225x == 2  ~ 2,
                                                   V161225x == 3  ~ 1,
                                                   V161225x == 7  ~ -3,
                                                   V161225x == 6  ~ -2,
                                                   V161225x == 5 ~ -1))


#calculation of Standard Deviation and mean for each party in  rising temperatures -------

anes_2020 %>% group_by(pid_chr) %>%
  summarise(sd_rising_temperatures  = sd(opinion_rising_temperatures , na.rm = TRUE))

anes_2020 %>% group_by(pid_chr) %>%
  summarise(mean_rising_temperatures  = mean(opinion_rising_temperatures , na.rm = TRUE))


anes_2016 %>% group_by(pid_chr) %>%
  summarise(sd_rising_temperatures  = sd(opinion_rising_temperatures , na.rm = TRUE))

anes_2016 %>% group_by(pid_chr) %>%
  summarise(mean_rising_temperatures  = mean(opinion_rising_temperatures , na.rm = TRUE))

#t-test----------

t.test(anes_2016$opinion_rising_temperatures, anes_2020$opinion_rising_temperatures)

#visualization--------
ggplot(data = anes_2016,
       mapping = aes(x=anes_2016$opinion_rising_temperatures, fill = anes_2016$pid_chr)) +
  geom_bar(position = "dodge")+scale_fill_manual(values=c('blue', 'green', 'red', 'white')) +theme_dark()

ggplot(data = anes_2020,
       mapping = aes(x=anes_2020$opinion_rising_temperatures, fill = anes_2020$pid_chr)) +
  geom_bar(position = "dodge")+scale_fill_manual(values=c('blue', 'green', 'red', 'white')) +theme_dark()


boxplot(anes_2016$opinion_rising_temperatures, anes_2020$opinion_rising_temperature, 
        names = c('2016', '2020'), col=c('red', 'blue'), main='Public oppinion in health care')


ggplot(anes_2016, aes(anes_2016$pid_chr, anes_2016$opinion_free_trade)) + 
  geom_boxplot(col=c('blue','green', 'red', 'white'))

ggplot(anes_2020, aes(anes_2020$pid_chr,anes_2020$opinion_rising_temperature)) + 
  geom_boxplot(col=c('blue','green', 'red', 'white'))




```



# FAVOR OR OPPOSE VOTE BY MAIL

```{r echo=FALSE}

  #create a variable for opinion_vote_mail for 2020
  anes_2020 <- anes_2020 %>% 
  mutate (opinion_vote_mail = case_when( V201356x == 4 ~ 0,
                                         V201356x == 1  ~ 3,  
                                         V201356x == 2  ~ 2,
                                         V201356x == 3  ~ 1,
                                         V201356x == 7  ~ -3,
                                         V201356x == 6  ~ -2,
                                         V201356x == 5 ~ -1)) 
#calculation of Standard Deviation and mean for each party in vote mail  trade (only for ANES-2020)----

anes_2020 %>% group_by(pid_chr) %>%
  summarise(sd_vote_mail = sd(opinion_vote_mail, na.rm = TRUE))



anes_2020 %>% group_by(pid_chr) %>%
  summarise(mean_vote_mail = mean(opinion_vote_mail, na.rm = TRUE))

ggplot(data = anes_2020,
       mapping = aes(x=anes_2020$opinion_vote_mail, fill = anes_2020$pid_chr)) +
  geom_bar(position = "dodge")+scale_fill_manual(values=c('blue', 'green', 'red', 'white')) +theme_dark()


ggplot(anes_2020, aes(anes_2020$pid_chr,anes_2020$opinion_vote_mail)) + 
  geom_boxplot(col=c('blue','green', 'red', 'white'))


boxplot(anes_2020$opinion_vote_mail, 
         col=c('blue'), main='Public oppinion vote by mail')
# End of codes -----
```

 **Thank you!**
